### 2022-11-15 NSF2022/Description/backgroundv3.tex

The field of self-organization encompasses a broad range of interesting
and challenging phenomena in engineering and science. For example,
understanding the organization of lipids and proteins in vesicles is
crucial to designing effective drug carriers and functional mimics of
cells~\cite{Marui2022IncreasedEO,
https://doi.org/10.1002/adma.202206288}. The COVID-19 pandemic has led
to a concerted effort to understand the mechanisms of RNA condensation
by structural proteins that control virus
structure~\cite{Kim2021SelfassembledMV}. Engineers have developed
superior techniques that rely on nanoscale capillary and van der Waals
forces to print high-resolution features on
devices~\cite{Zeng20223DprintedMT}. Finally, researchers design
efficient algorithms for constructing interaction kernels in the to
study of agent-based systems like swarms and
flocking~\cite{Lu2019NonparametricIO, Tadmor2021OnTM}.

@@

When fluids are involved, tracking self-organization becomes difficult
due to the deformability of suspended media in the presence of non-local
hydrodynamic interactions. The phase field method, a prominent approach
in this area, introduces a scalar-valued phase field function for the
volume fraction of amphiphilic lipids or polymers in solvent, for
example. Sometimes called the diffuse interface method, one devises
energy functionals involving the gradient and well-potentials of the
phase field, and possibly convolution kernels, to characterize the self-
and solvent-interaction of the suspended
media~\cite{Promislow2022UndulatedBI, C9SM01983A,doi:10.1063/5.0009734,
LiAn-Chang16, Choksi2003OnTD}. Because the fluid interfaces are given by
level-sets of a smooth function, they can have almost arbitrary topology
e.g., free ends, junctions~\cite{Promislow2017ExistenceBA,
Promislow2022UndulatedBI}, the evolution equations can be studied by
well-established variational and analytical
tools~\cite{Gavish2011CurvatureDF, Dai2019WeakSF, Dai2015CompetitiveGE,
Dai2022GeometricEO, Dai2020MinimizersFT, Dai2013GeometricEO}, and the
governing equations are amenable to numerical
approximation~\cite{Christlieb2020BenchmarkCO,
Christlieb2019CompetitionAC}. One of the downsides, though, is that
granular details like molecular heterogeneity and orientations that may
be crucial to the physics at mesoscopic scales \cite{Midya2022}, are blurred
by the model. Moreover, numerical simulations of phase field
models typically discretize the entire fluid domain, and apply cutoff
boundary conditions when periodic domains are not under consideration.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{figures/Background/Domain.pdf}
%    \includegraphics[width=0.45\textwidth]{figures/SpecificAim1/Domain.jpg}
%    \includegraphics[width=0.45\textwidth]{figures/SpecificAim1/3Particles.jpg}
%    \includegraphics[keepaspectratio,height=2.7cm]{figures/SpecificAim1/Domain.jpg}
%    \includegraphics[keepaspectratio,height=2.7cm]{figures/SpecificAim1/3Particles.jpg}
  \end{center}
  \caption{\label{fig:flow_map} \footnotesize (a) Granules make a
  superstructure e.g., a vesicle. The panel shows the moving domain
  $\Omega(t)$, granules with centers $\aa_i$ and directors $\dd_i$, and
  the unit inward normal $\nnu$. Panels (b), (c), (d) show the interplay
  between the phase field $\phi$ (colormap) and granule configuration.
  Red is for $\phi = 1$ and blue for $\phi = 0$; $W(\phi) =
  \tfrac{1}{2}\phi^2$. The white arrows are the fluid velocity field.
  The panels reach an equilibrium state in panel (d).}
\end{figure}

@@

To determine how the granules move through space, we look for rigid body
transformations
\begin{align}
\label{eq:RBT}
  \FF_i(\XX,t) = R_i(t)(\XX - \aa_i(0)) + \aa_i(t),\quad \XX \in U_i(0),
\end{align}
where $i = 1,\ldots,N_b,$ $R_i(t)$ is an $n
\times n$ rotation matrix, $\aa_i(t)$ are the granule centers, and $\XX$ is a
reference coordinate.
The dynamics come from the following system of semi-linear, elliptic boundary value problems:
\begin{alignat}{4}
  \label{eqn:stokes} 
  -\mu\Delta \uu + \nabla p &= \mathbf{0}, 
  \quad \nabla \cdot \uu = 0, &&\xx \in \Omega(t), \\
  \label{eqn:phase}
  -\epsilon^2 \Delta \phi + W'(\phi) &= 0, &&\xx \in \Omega(t),\\
  \label{eqn:noslip}        
  \frac{d\FF_i}{dt}(\XX,t) & = \uu(\xx) = 
    \vv_i + \oomega_i \times (\xx - \aa_i), 
  \quad &&\xx \in \Gamma_i(t),\\
  \label{eqn:material}
  \phi(\xx) &= h_i(\XX),  &&\xx \in \Gamma_i(t),
\end{alignat}
for $i=1,\ldots,N_b$ and where $\xx = \FF_i(\XX,t)$. Finally,
\begin{align}
\label{eqn:stressbalance}
\int_{\Gamma_i} \left(\sigma  + T_i\right)\nnu \,\dif s = \mathbf{0},\quad
\int_{\Gamma_i} (\xx - \aa_i)\times \left(\sigma + T_i\right) \nnu
  \,\dif s = \mathbf{0},
\end{align}
where
\begin{align}
\label{eqn:hydro_stress}
\sigma = \mu(\nabla \uu + \nabla \uu^T) - pI,\quad 
T_i = \gamma\left[\epsilon^{-1} W(\phi)I
  + \epsilon\left(\tfrac{1}{2}|\nabla \phi|^2I - \nabla \phi \nabla
  \phi^T\right)\right]
\end{align}
are the hydrodynamic and phase field stresses, respectively.
Equations~\eqref{eqn:stokes} are the Stokes equations describing the
velocity $\uu$ and pressure $p$ of the incompressible fluid region with
viscosity $\mu$. Equation~\eqref{eqn:phase} describes the transitions
of the scalar order parameter $\phi$ with a decay length $\epsilon$.
The scalar function $W(\phi) : \RR \to \RR$ is a nonnegative
well-potential with local minima $\phi_0, \phi_1$, etc., representing
metastable states of the solvent. In the single-well case, $W(\phi) =
\frac{1}{2}\phi^2$ has the minimum $\phi_0 = 0$ and the phase field
equation~\eqref{eqn:phase} becomes the \emph{linear} screened Laplace
equation. In the double-well case e.g, $W(\phi) =
\frac{1}{4}(\phi^2-1)^2+cx$, $c > 0$, the phase field
equation~\eqref{eqn:phase} takes the form of an Allen-Cahn equation with
local minima $\phi_0 < 0 < \phi_1$. Section \ref{sec:specificaim1}
discusses certain modifications needed for~\eqref{eqn:phase} in the
nonlinear case.

@@

\begin{wrapfigure}[10]{l}{0.5\textwidth}
\vspace{-5pt}
\includegraphics[width=0.5\textwidth]{figures/PreliminaryWork/TankTreading.jpg}
\vspace{-20pt}
\caption{\label{fig:JPv_linearshear} \footnotesize Granular vesicles
  undergo tank-treading in background shear flow.}
\end{wrapfigure}
%
A striking feature of our hydrodynamic simulations is their fidelity to
physics. The model contains a few parameters; granule radius $c$, decay
length $\epsilon$, surface tension $\gamma$. When these are set to
realistic values \cite{Fu2018_SIAM, ErLjCl89, Lin2005, Parsegian,
Israelachvili80, GarciaSaez, KUZMIN2005, Petelska2012,Jackson2016}, we
quantitatively recover well-known elastic moduli for
single-component lipid bilayers \cite{Nagle17, Nagle17-2,
LeVeWa14,NAGLE2000159}. Such agreement is a validation that
equations~\eqref{eq:RBT}--\eqref{eqn:stressbalance} correctly capture
the long-range interactions between amphiphilic granules (such as lipid
macromolecules) surfaces.
\begin{wrapfigure}[13]{l}{0.475\textwidth}
\vspace{-0pt}
\includegraphics[width=0.475\textwidth]{figures/PreliminaryWork/Rupture.jpg}
\caption{\label{fig:JPv_rupture} \footnotesize Rupture of a granular
  vesicle under large shear rates.}
\end{wrapfigure}

@@

\begin{wrapfigure}[16]{r}{4in} 
  \includegraphics[width=4in]{figures/Background/coarsening.pdf}
  \vspace{-25pt}
  \caption{\label{fig:coarsening} \footnotesize Energy $E$ in
  equation~(\ref{eq:energy_law}) of a suspension of amphiphilic granules
  versus time $t$. Under the moving domain problem
  \eqref{eq:RBT}--\eqref{eqn:hydro_stress} with amphiphilic boundary
  condition, the granules self-organize into vesicle bilayers. External
  hydrodynamic effects are introduced to study elastic properties of
  the granules assembly~\cite{Fu2018_SIAM,FuQuRyYo22}.}
\end{wrapfigure}

>>>>

The field of self-organization encompasses a broad range of interesting
and challenging phenomena in engineering and science. For example,
understanding the organization of lipids and proteins in vesicles is
crucial to designing effective drug carriers and functional mimics of
cells~\cite{Marui2022IncreasedEO,
https://doi.org/10.1002/adma.202206288}. The COVID-19 pandemic has led
to a concerted effort to understand the mechanisms of RNA condensation
by structural proteins that control virus
structure~\cite{Kim2021SelfassembledMV}. Engineers develop superior
techniques that rely on nanoscale capillary and van der Waals forces to
print high-resolution features on devices~\cite{Zeng20223DprintedMT}.
Finally, researchers design efficient algorithms for constructing
interaction kernels in the to study of agent-based systems like swarms
and flocking~\cite{Lu2019NonparametricIO, Tadmor2021OnTM}.

@@

When fluids are involved, tracking self-organization becomes difficult
due to the deformability of suspended media in the presence of non-local
hydrodynamic interactions. The phase field method, a prominent approach
in this area, introduces a scalar-valued phase field function for the
volume fraction of amphiphilic lipids or polymers in solvent, for
example. Sometimes called the diffuse interface method, one devises
energy functionals involving the gradient and well-potentials of the
phase field, and possibly convolution kernels, to characterize the self-
and solvent-interaction of the suspended
media~\cite{Promislow2022UndulatedBI, C9SM01983A,doi:10.1063/5.0009734,
LiAn-Chang16, Choksi2003OnTD}. Because the fluid interfaces are given by
level-sets of a smooth function, they can have almost arbitrary topology
such as free ends and junctions~\cite{Promislow2017ExistenceBA,
Promislow2022UndulatedBI}, the evolution equations can be studied by
well-established variational and analytical
tools~\cite{Gavish2011CurvatureDF, Dai2019WeakSF, Dai2015CompetitiveGE,
Dai2022GeometricEO, Dai2020MinimizersFT, Dai2013GeometricEO}, and the
governing equations are amenable to numerical
approximation~\cite{Christlieb2020BenchmarkCO,
Christlieb2019CompetitionAC}. One of the downsides, though, is that
granular details like molecular heterogeneity and orientations that may
be crucial to the physics at mesoscopic scales \cite{Midya2022}, are blurred
by the model. Moreover, numerical simulations of phase field
models typically discretize the entire fluid domain, and apply cutoff
boundary conditions when periodic domains are not under consideration.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{figures/Background/Domain.pdf}
%    \includegraphics[width=0.45\textwidth]{figures/SpecificAim1/Domain.jpg}
%    \includegraphics[width=0.45\textwidth]{figures/SpecificAim1/3Particles.jpg}
%    \includegraphics[keepaspectratio,height=2.7cm]{figures/SpecificAim1/Domain.jpg}
%    \includegraphics[keepaspectratio,height=2.7cm]{figures/SpecificAim1/3Particles.jpg}
  \end{center}
  \caption{\label{fig:flow_map} \footnotesize (a) Granules make a
  superstructure e.g., a vesicle. The panel shows the moving domain
  $\Omega(t)$, granules with centers $\aa_i$ and directors $\dd_i$ with
  $|\dd_i| = 1$, and the unit inward normal $\nnu$. Panels (b), (c), (d)
  show the interplay between the phase field $\phi$ (colormap) and
  granule configuration. Red is for $\phi = 1$ and blue for $\phi = 0$;
  $W(\phi) = \tfrac{1}{2}\phi^2$. The white arrows are the fluid
  velocity field. The panels reach an equilibrium state in panel (d).}
\end{figure}

@@

To determine how the granules move through space, we look for rigid body
transformations
\begin{align}
\label{eq:RBT}
  \FF_i(\XX,t) = R_i(t)(\XX - \aa_i(0)) + \aa_i(t),\quad \XX \in U_i(0),
\end{align}
where $i = 1,\ldots,N_b,$ $R_i(t)$ is an $n
\times n$ rotation matrix, $\aa_i(t)$ are the granule centers, and $\XX$ is a
reference coordinate.
The dynamics come from the following system of semi-linear, elliptic boundary value problems:
\begin{alignat}{4}
  \label{eqn:stokes} 
  -\mu\Delta \uu + \nabla p &= \mathbf{0}, 
  \quad \nabla \cdot \uu = 0, &&\xx \in \Omega(t), \\
  \label{eqn:phase}
  -\epsilon^2 \Delta \phi + W'(\phi) &= 0, &&\xx \in \Omega(t),\\
  \label{eqn:noslip}        
  \frac{d\FF_i}{dt}(\XX,t) & = \uu(\xx) = 
    \vv_i + \oomega_i \times (\xx - \aa_i), 
  \quad &&\xx \in \Gamma_i(t),\\
  \label{eqn:material}
  \phi(\xx) &= h_i(\XX),  &&\xx \in \Gamma_i(t),
\end{alignat}
for $i=1,\ldots,N_b$ and where $\xx = \FF_i(\XX,t)$. Finally,
\begin{align}
\label{eqn:stressbalance}
\int_{\Gamma_i} \left(\sigma  + T_i\right)\nnu \,\dif s = \mathbf{0},\quad
\int_{\Gamma_i} (\xx - \aa_i)\times \left(\sigma + T_i\right) \nnu
  \,\dif s = \mathbf{0},
\end{align}
where
\begin{align}
\label{eqn:hydro_stress}
\sigma = \mu(\nabla \uu + \nabla \uu^T) - pI,\quad 
T_i = \gamma\left(\frac{1}{\epsilon}W(\phi)I + \frac{\epsilon}{2}|\nabla \phi|^2 - \epsilon \nabla \phi \nabla \phi^T\right)
%T_i = \gamma\left[\epsilon^{-1} W(\phi)I
%  + \epsilon\left(\tfrac{1}{2}|\nabla \phi|^2I - \nabla \phi \nabla
%  \phi^T\right)\right]
\end{align}
are the hydrodynamic and phase field stresses, respectively.
Equations~\eqref{eqn:stokes} are the Stokes equations describing the
velocity $\uu$ and pressure $p$ of the incompressible fluid region with
viscosity $\mu$. Equation~\eqref{eqn:phase} describes the transitions
of the scalar order parameter $\phi$ with a decay length $\epsilon$.
The scalar function $W(\phi) : \RR \to \RR$ is a nonnegative
well-potential with local minima $\phi_0, \phi_1$, etc., representing
metastable states of the solvent. In the single-well case, $W(\phi) =
\frac{1}{2}\phi^2$ has the minimum $\phi_0 = 0$ and the phase field
equation~\eqref{eqn:phase} becomes the \emph{linear} screened Laplace
equation. In the double-well case e.g, $W(\phi) =
\frac{1}{4}(\phi^2-1)^2+cx$, $c > 0$, the phase field
equation~\eqref{eqn:phase} takes the form of an Allen-Cahn equation with
local minima $\phi_0 < 0 < \phi_1$. Section \ref{sec:specificaim1}
discusses certain modifications needed for~\eqref{eqn:phase} in the
nonlinear case.

@@

\begin{wrapfigure}[10]{l}{0.5\textwidth}
\vspace{-5pt}
\includegraphics[width=0.5\textwidth]{figures/PreliminaryWork/TankTreading.jpg}
\vspace{-20pt}
\caption{\label{fig:JPv_linearshear} \footnotesize Granular vesicles
  undergo tank-treading in background shear flow.}
\end{wrapfigure}
%
A striking feature of our hydrodynamic simulations is their fidelity to
physics. The model contains a few parameters; granule radius $c$, decay
length $\epsilon$, surface tension $\gamma$. When these are set to
realistic values \cite{Fu2018_SIAM, ErLjCl89, Lin2005, Parsegian,
Israelachvili80, GarciaSaez, KUZMIN2005, Petelska2012,Jackson2016}, we
quantitatively recover well-known elastic moduli for
single-component lipid bilayers \cite{Nagle17, Nagle17-2,
LeVeWa14,NAGLE2000159}. Such agreement is a validation that
equations~\eqref{eq:RBT}--\eqref{eqn:stressbalance} correctly capture
the long-range interactions between amphiphilic granules (such as lipid
macromolecules) surfaces.
\begin{wrapfigure}[13]{l}{0.475\textwidth}
\vspace{-0pt}
\includegraphics[width=0.475\textwidth]{figures/PreliminaryWork/Rupture.jpg}
\vspace{-25pt}
\caption{\label{fig:JPv_rupture} \footnotesize Rupture of a granular
  vesicle under large shear rates.}
\end{wrapfigure}

@@

\begin{wrapfigure}[16]{r}{4in} 
  \vspace{-10pt}
  \includegraphics[width=4in]{figures/Background/coarsening.pdf}
  \vspace{-25pt}
  \caption{\label{fig:coarsening} \footnotesize Energy $E$ in
  equation~(\ref{eq:energy_law}) of a suspension of amphiphilic granules
  versus time $t$. Under the moving domain problem
  \eqref{eq:RBT}--\eqref{eqn:hydro_stress} with amphiphilic boundary
  condition, the granules self-organize into vesicle bilayers. External
  hydrodynamic effects are introduced to study elastic properties of
  the granules assembly~\cite{Fu2018_SIAM,FuQuRyYo22}.}
\end{wrapfigure}

### 2022-11-15 NSF2022/Description/specificaim2.tex
...

As an illustration, the problem of finding the global electrostatic
potential $u$ with surface charge density $\sigma$ is a special case of
\eqref{eqn:conv}, where $G(\xx - \yy) = (4\pi \epsilon_0)^{-1}|\xx -
\yy|^{-1}$ in three dimensions ($-(2\pi \epsilon_0)^{-1}\log|\xx - \yy|$
in two dimensions) with $\epsilon_0$ the permittivity of vacuum.  In
this context, equation~\eqref{eqn:bie} is analogous to Gauss's Law.

@@

Having demonstrated that the inexact Newton method converges in one
dimension, it will be coupled with IE formulations to
solve~\eqref{eqn:phaseRoot} in two and three dimensions. The solution of
$-\epsilon^2 \Delta v + \beta v = F[\phi^N]$ can be written
as $v = v^P + v^H$ where $v^P$ is a volume potential and $v^H$ is a
homogeneous solution of the PDE, which is represented as a layer
potential, and is used to satisfy the boundary condition. The volume
potential is
\begin{align*}
  v^{P}(\xx) = \int_{\Omega(t)} G(\xx,\yy) F[\phi^N](\yy)\, d\yy, 
    \quad \xx \in \Omega(t),
\end{align*}
where $G$ is the fundamental solution of $-\epsilon^2 \Delta +
\beta$. Traditionally, volume potentials are more challenging
to compute than layer potentials, but new methods have recently been
developed that allow them to be computed accurately and
efficiently~\cite{fry-leh-tor2018, and-zhu-vee2022}. This approach of
combining volume and layer potentials has been successfully used by PI
Quaife~\cite{kro-qua2010} and others~\cite{fry-kro-tor2019,
kli-ask-kro2020, che-hua-lei2005} in other contexts, such as time
stepping.

>>>>

As an illustration, the problem of finding the global electrostatic
potential $u$ with surface charge density $\sigma$ is a special case of
\eqref{eqn:conv}, where $G(\xx - \yy)$  takes the form
$-(2\pi \epsilon_0)^{-1}\log|\xx - \yy|$
in two dimensions and 
$(4\pi \epsilon_0)^{-1}|\xx -\yy|^{-1}$ in three dimensions
with $\epsilon_0$ the permittivity of vacuum.  In
this context, equation~\eqref{eqn:bie} is analogous to Gauss's Law.

@@

Having demonstrated that the inexact Newton method converges in one
dimension, it will be coupled with IE formulations to
solve~\eqref{eqn:phaseRoot} in two and three dimensions. The solution of
$-\epsilon^2 \Delta v + \beta v = F[\phi^N]$ can be written
as $v = v^P + v^H$ where $v^P$ is a volume potential and $v^H$ is a
homogeneous solution of the PDE, which is represented as a layer
potential, and is used to satisfy the boundary condition. The volume
potential is
\begin{align}
  v^{P}(\xx) = \int_{\Omega(t)} G(\xx,\yy) F[\phi^N](\yy)\, d\yy, 
    \quad \xx \in \Omega(t),
\end{align}
where $G$ is the fundamental solution of $-\epsilon^2 \Delta +
\beta$. Traditionally, volume potentials are more challenging
to compute than layer potentials, but new methods have recently been
developed that allow them to be computed accurately and
efficiently~\cite{fry-leh-tor2018, and-zhu-vee2022}. This approach of
combining volume and layer potentials has been successfully used by PI
Quaife~\cite{kro-qua2010} and others~\cite{fry-kro-tor2019,
kli-ask-kro2020, che-hua-lei2005} in other contexts, such as time
stepping.
